---
title: "Lab 1"
author: "Ksenia Gordeeva"
date: "4/9/2022"
output:
  html_document:
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(rio)
library(here)
library(ggplot2)
```

### Data Structures


```{r}
set.seed(123) # Set random number generation

m <- lm(mpg ~ hp + cyl + drat, mtcars)

l <- list(a = list(m = matrix(1:12, ncol = 3),
              	   v = 1:7),
          b = data.frame(student = 1:15,
                         score = rnorm(15, 100, 10))) 
```

1. From the model results:
	* Extract the qr tolerance (`tol`).
```{r}
#str(m)
#Extract the qr tolerance
m[["qr"]][["tol"]]

#Alternatively: 
#m$qr$tol
```

	* Extract the term labels from the model.
```{r}
#Term Labels
term_attributes <- attr(m$model, "terms")
attr(term_attributes, "term.labels")

#Alternatively:
#attr(m[["terms"]], "term.labels")
```

2. From the list
	* Use at least two different methods to extract `m`.
```{r}
#1
l[["a"]][["m"]]

#2
l$a$m
```

	* Extract the third column of `m`. Maintain the dimensions (matrix structure).
```{r}
l[["a"]][["m"]][ ,3, drop = FALSE]
```

	* Extract the score for student 7.
```{r}
l$b[7, ]

#ALternatively: 
#l[["b"]][7, ]
```

### Loops
3. Read in the `insurance_coverage.csv` file, which is in the data folder in the repo. Split the data file by `age_bucket` and `sex_name`.

```{r}
ins_cov <- read_csv(here("~/Documents/Uoregon/spring 2022/c3-fp-2022/data/insurance_coverage.csv"))

#by_age <- split(ins_cov, ins_cov$age_bucket)

#by_sexname <- split(ins_cov, ins_cov$sex_name)

by_age_sex <- split(ins_cov, 
                    list(ins_cov$age_bucket, ins_cov$sex_name))
```

4. Use a `for` loop to calculate the mean health care population (`hc_pop`) for each `age_bucket`/`sex_name` combination.
```{r}
hc_age_sex <- vector("list", length(by_age_sex))

for(i in seq_along(hc_age_sex)) {
	hc_age_sex[i] <- mean(by_age_sex[[i]]$hc_pop, na.rm = TRUE)
	names(hc_age_sex)[i] <- names(by_age_sex)[i]
}

```
5. Replicate this calculation with `lapply`, `sapply`, and `vapply`

```{r}
## lapply
hc_age_sex2 <- lapply(by_age_sex, function(df) mean(df$hc_pop, na.rm = TRUE))

## sapply
hc_age_sex3 <- sapply(by_age_sex, function(df) mean(df$hc_pop, na.rm = TRUE))

## vapply
hc_age_sex4 <- vapply(by_age_sex, function(df) mean(df$hc_pop, na.rm = TRUE),
                      FUN.VALUE = double(1))
```
6. Produce separate plots showing the change in `hc_pop` from 2014 to 2015 for each `age_bucket`/`sex_name` combination (with a single loop). Set the `hc_pop` axis so the limits equal the overall minimum and maximum values for `hc_pop`.

```{r}
min <- min(ins_cov$hc_pop)
max <- max(ins_cov$hc_pop)

theme_set(theme_minimal())

plots <- lapply(by_age_sex, function(df) {
	ggplot(df, aes(x = as.factor(year), y = hc_pop)) +
		geom_col(alpha = .6) +
    coord_cartesian(ylim = c(min, max)) +
    labs(title = names(by_age_sex)[i],
         x = "Year",
         y = "Health Care Population")
    #theme_minimal()
})

plots
```
7. Use a `for` loop to save the plots to a folder on your computer
```{r}
dir.create(here::here("plots"))
filenames <- paste0(here::here("plots", "hc_pop"), names(by_age_sex), ".png")

for(i in seq_along(plots)) {
	ggsave(filenames[i],
	       plots[[i]],
	       device = "png",
	       width = 6.5,
	       height = 8)
}
```



